High-Dimensional Regularized Discriminant Analysis

### 2014-04-02 reports/paper-hdrda.Rnw
Much needed edits applied.

\citeauthor{Friedman:1989tm} proposed the popular regularized discriminant
analysis (RDA) classifier that utilizes a biased covariance-matrix estimator to
partially pool the sample covariance matrices from linear and quadratic
disriminant analysis before shrinking the resulting estimator towards a scaled
identity matrix. The RDA classifier's two tuning parameters are typically
estimated via a computationally burdensome cross-validation procedure from a
grid of candidate values. We reformulate the RDA classifier in the
high-dimensional setting and then show that the classification decision rule is
equivalent to one in a subspace having a much lower dimension. As a result, the
dimension reduction involved yields a substantial reduction in computation
during model selection. Our parameterization also offers interpretability
previously lacking with the RDA classifier. We demonstrate with three artificial
and four real high-dimensional data sets that our proposed classifier is often
superior to several recently proposed sparse and regularized classifiers in
terms of classification accuracy. Finally, we provide an implementation of our
proposed classifier in the {\tt sparsediscrim} R package, available on CRAN.

@@

We reparameterize the \emph{RDA} classifier similar to \cite{Hastie:2008dt} and
\cite{Halbe:2007ea} and refer to our parameterization as the high-dimensional
\emph{RDA} (\emph{HDRDA}) classifier. First, we show the pooling parameter in
the \emph{HDRDA} classifier determines the contribution of each training
observation to the estimation of each class covariance matrix, enabling
interpretability that has been previously lacking with the \emph{RDA} classifier
\citep{Bensmail:1996fw}. Furthermore, the \emph{HDRDA} classifier includes a
large family of shrinkage methods, such as those from \cite{Srivastava:2007ww},
\cite{Rao:1971ul}, and other methods studied by \cite{Ramey:2013ji} and
\cite{Xu:2009fl}. Next, applying a reasoning similar to \cite{Ye:2006tq}, we
show that the matrix operations corresponding to the null space of the pooled
sample covariance matrix are redundant and can be discarded from the
\emph{HDRDA} decision rule without loss of classificatory information. Moreover,
because the \emph{HDRDA} decision rule is invariant to adjustments to the zero
eigenvalues, the decision rule in the original feature space is equivalent to a
decision rule in a much lower dimension (i.e., the intrinsic dimension), such
that the matrix inverses and determinants of relatively small matrices can be
rapidly computed. Thus, the dimension reduction inherent in the \emph{HDRDA}
classifier yields a substantial reduction in computation during model selection.

@@

In discriminant analysis we wish to assign correctly an unlabeled vector $\bm x
\in \mathbb{R}_{p \times 1}$ to one of $K$ unique, known classes by constructing
a classifier from $N$ training observations. Let $\bm x_i = (x_{i1}, \ldots,
x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots,
N)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$. We
assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution $p(\bm
x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$
is the probability density function of the $k$th class and $p(\omega_k)$ is the
prior probability of class membership of the $k$th class. We further assume that
each of $p(\omega_k)$ is equal, $k = 1, \ldots, K$.

@@

The smallest eigenvalues and the directions associated with their eigenvectors
can highly influence \eqref{eq:qda}. In fact, the eigenvalues of $\widehat{ \bm
  \Sigma}_k$ are well-known to be biased if $p \ge n_k$ such that the smallest
eigenvalues are underestimated \citep{Seber:2004uh}. Moreover, if $p > n_k$,
then rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that at least $p -
n_k$ eigenvalues of $\widehat{\bm \Sigma}_k$ are zero. Furthermore, although
more feature information is available to discriminate among the $K$ classes, if
$p > n_k$, \eqref{eq:qda} is incalculable because $\widehat{\bm \Sigma}_k^{-1}$
does not exist.

@@

Here, we define the \emph{HDRDA} classifier by first formulating the
covariance-matrix estimator $\widehat{\bm \Sigma}_k(\lambda)$ and demonstrating
its clear interpretation as a linear combination of the crossproducts of the
training observations centered by their respective class sample means. We define
the convex combination
\begin{align}
  \widehat{\bm \Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm \Sigma}_k + \lambda
  \widehat{\bm \Sigma}, \quad k = 1, \ldots, K,\label{eq:sig-lambda-alternative}
\end{align}
where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. By rewriting
\eqref{eq:sig-lambda-alternative} in terms of the observations $ \bm x_i$ $(i =
1, \ldots, N)$, each centered by its class sample mean, we attain a clear
interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is, from
\eqref{eq:pooled-cov} we have
\begin{align}
	\widehat{\bm \Sigma}_k(\lambda)
  &= \left( 1 - \lambda + \frac{\lambda n_k}{N} \right) \widehat{\bm \Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}}^K n_{k'} \widehat{\bm \Sigma}_{k'} \nonumber \\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N
  I(y_i = k) \bm x_i \bm x_i^{T} +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k)
  \bm x_i \bm x_i^{T} \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i^{T},\label{eq:sig-lambda-alternative2}
\end{align}
where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$. We
can interpret \eqref{eq:sig-lambda-alternative2} as a covariance-matrix
estimator that borrows from $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov} to
better estimate $ \bm \Sigma_k$. From \eqref{eq:sig-lambda-alternative2},
$\lambda$ weights the contribution of each of the $N$ observations in estimating
$\bm \Sigma_k$. Hence, for $0 < \lambda \le 1$, we estimate $\bm \Sigma_k$ with
$N$ rather than $n_k$ observations.\footnote{TODO: Add more to the
  interpretation here and reference Figure \ref{fig:hdrda-weights}.}

\begin{figure}
<<rda_weights>>=
n <- c(50, 25, 15, 10)
N <- sum(n)
lambda <- seq(0, 1, length = 1000)

rda_weights <- sapply(seq_along(n), function(k) {
      lambda / N + (1 - lambda) / n[k]
      })
rda_weights <- data.frame(lambda = lambda, rda_weights)
rda_weights <- melt(rda_weights, id = 'lambda')
colnames(rda_weights) <- c("lambda", "Class", "Weight")
levels(rda_weights$Class) <- seq_along(n)

p <- ggplot(rda_weights, aes(lambda, Weight, linetype = Class, color = Class))
p <- p + geom_line(size = 1)
p <- p + theme_bw() + xlab(expression(lambda))
p <- p + theme(axis.text = element_text(size = 10))
p <- p + theme(strip.text.x = element_text(size = 16))
p <- p + theme(legend.title = element_text(size = 16, face = "bold"))
p <- p + theme(legend.text = element_text(size = 15))
p <- p + theme(legend.key.size = unit(1, "cm"))
p
@ 
\caption{Weights $c_{ik}(\lambda)$ of \emph{RDA} classifier as a function of the
  pooling parameter $\lambda$ when $n_1 = 50$, $n_2 = 25$, $n_3 = 15$, and $n_4$
  = 10. TODO: Scale down figure.}
\label{fig:hdrda-weights}
\end{figure}

Below, we will show that the pooling
operation is advantageous in increasing the rank of each $\widehat{\bm
  \Sigma}_k(\lambda)$ from rank$(\widehat{\bm \Sigma}_k)$ to rank$(\widehat{\bm
  \Sigma})$ for $0 < \lambda \le 1$. Notice that if $\lambda = 0$, then the
observations from the remaining $K - 1$ classes do not contribute to the
estimation of $\bm \Sigma_k$, corresponding to $\widehat{\bm
  \Sigma}_k$. Furthermore, if $\lambda = 1$, the weights in
\eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to
$\widehat{\bm \Sigma}$. For brevity, when $\lambda = 1$, we define $\bm X =
      [\sqrt{c_{1k}(1)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} \bm x_N^{T}]^{T}$
      such that $\widehat{\bm \Sigma} = \bm X^{T} \bm X$. Similarly, for
      $\lambda = 0$, we define $\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1^{T}, \ldots,
        \sqrt{c_{Nk}(0)} \bm x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma}_k =
      \bm X_k^{T} \bm X_k$. In Figure \ref{fig:hdrda-contours} we plot the
      contours of five multivariate normal populations for $\lambda = 0$ with
      unequal covariance matrices. As $\lambda$ approaches 1, the contours
      become more similar, resulting in identical contours for $\lambda = 1$.
\begin{figure}
<<hdrda_contours, fig.width=8, fig.height=8, out.width='.33\\linewidth', fig.show='hold'>>=
rda_contours(lambda = 0)
rda_contours(lambda = 1/4)
rda_contours(lambda = 3/4)
rda_contours(lambda = 1)
@ 
\caption{Contours of five multivariate normal populations as a function of the
  pooling parameter $\lambda$.}
\label{fig:hdrda-contours}
\end{figure}
As we have discussed above, several eigenvalue adjustment methods have been
proposed to increase eigenvalues (approximately) equal to 0. To further improve
the estimation of $\bm \Sigma_k$ and to stabilize the estimator's inverse, we
define the eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
\begin{align}
	\tilde{\bm \Sigma}_k = \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:hdrda-cov}
\end{align}
where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
constant. Thus, the \emph{pooling} parameter $\lambda$ controls the amount
borrowed from $\widehat{\bm \Sigma}$ to estimate $\bm \Sigma_k$, and the
\emph{shrinkage} parameter $\gamma$ determines the degree of eigenvalue
shrinkage. The choice of $\alpha_k$ allows for a flexible formulation of
covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1, \ldots,
K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}. Similarly,
if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form similar to the
\emph{RDA} classifier from \cite{Friedman:1989tm}. Now, substituting
\eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the \emph{HDRDA} classifier
\begin{align}
	D_{HDRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)^{T}\tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k|. \label{eq:hdrda}
\end{align}
For $\gamma > 0$, $\tilde{\bm \Sigma}_k$ is nonsingular such that $\tilde{\bm
  \Sigma}_k^{-1}$ can be substituted for $\tilde{\bm \Sigma}_k^{+}$ in
\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\bm \Sigma}_k|$
equal to the product of the positive eigenvalues of $\tilde{\bm
  \Sigma}_k$. Following \cite{Friedman:1989tm}, we select $\lambda$ and $\gamma$
from a grid of candidate models via cross-validation \citep{Hastie:2008dt}. We
provide an implementation of \eqref{eq:hdrda} in the {\tt hdrda} function in the
{\tt sparsediscrim} R package, available on CRAN.

@@

To further improve the estimation of $\bm \Sigma_k$ and to stabilize the inverse
of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased
covariance-matrix estimator
\begin{align}
          \widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm
            \Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm
              \Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
          \end{align}
where $\gamma \in [0, 1]$ is a regularization parameter that controls the
shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
of the eigenvalues of \eqref{eq:sig-lambda}. Finally, the RDA classifier is
established by substituting \eqref{eq:sig-rda} into
\eqref{eq:qda}.

\subsection{Properties of the HDRDA Classifier}
\label{sec:hdrda-properties}

@@

Thus, by incorporating each $\bm x_i$ into the estimation of $\bm \Sigma_k$, we
increase the rank of $\widehat{\bm \Sigma}_k(\lambda)$ to $q$ if $\lambda \ne
0$. The results is that the \emph{HDRDA} classifier has the advantage over
\eqref{eq:qda} in that the rank$(\widehat{\bm \Sigma}_k) \le n_k$. Next, we
provide an essential result that enables us to prove that \eqref{eq:hdrda} is
invariant to adjustments to the eigenvalues of $\tilde{ \bm \Sigma}_k$
corresponding to the null space of $\widehat{\bm \Sigma}$ for any unlabeled
observation $\bm x \in \mathbb{R}_{p \times 1}$.

@@

\begin{proposition}\label{proposition:hdrda-W_k}
Let $\bm W_k$ be defined as above, $k = 1, \ldots, K$. Then, $|\bm W_k| = |
\bm \Gamma_k| |\bm Q_k|$ and
	\begin{align}
		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - \alpha_k(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T} \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},\label{eq:W_k_inv}
    \intertext{where}
    \bm Q_k &= \bm I_{n_k} + \alpha_k(1 - \lambda) \bm X_k \bm U_1 \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T}\label{eq:Q_k}
    \intertext{and}
    \bm \Gamma_k &= \alpha_k \lambda \bm D_q + \gamma \bm I_q\label{eq:Gamma_k}.
	\end{align}
\end{proposition}
\begin{proof}
To calculate $|\bm W_k|$, we apply Theorem 18.1.1 from \cite{Harville:2008wja},
which states that $|\bm A + \bm B \bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} +
\bm C \bm A^{-1} \bm B|$. Thus, setting $\bm A = \bm \Gamma_k$, $\bm B =
\alpha_k (1 - \lambda) \bm U_1^{T} \bm X_k^{T}$, $\bm T = \bm I_{n_k}$, and $\bm
C = \bm X_k \bm U_1$, we have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$. Also,
\eqref{eq:W_k_inv} follows from the well-known Sherman-Woodbury formula
\citep[Theorem 18.2.8]{Harville:2008wja}.
\end{proof}

@@

        Compute $(\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k) + \log | \bm W_k |$ \tcp*[r]{TODO: Use matrix form for test observations}
      }
      Classify test observations $\bm X_{test}$ using \eqref{eq:hdrda-decomposed}

@@

A primary contributing factor to the efficiency of Algorithm
\ref{alg:hdrda_model_selection} is our usage of the compact singular value
decomposition (SVD) (sometimes called the reduced or fast SVD). Rather than
computing the eigenvalue decomposition of $\widehat{\bm \Sigma}$ to obtain $\bm
U_1$, we instead obtain $\bm U_1$ by computing the eigendecomposition of a much
smaller $N \times N$ matrix when $p \gg N$
\cite[Chapter~18.3.5]{Hastie:2008dt}. Applying the SVD, we decompose $\bm X_c =
\bm M \bm \Delta \bm U^{T}$, where $\bm M \in \mathbb{R}_{N \times p}$ is
orthogonal, $\bm \Delta \in \mathbb{R}_{p \times p}^{\ge}$ is a diagonal matrix
consisting of the singular values of $\bm X_c$, and $\bm U \in \mathbb{R}_{p
  \times p}$ is orthogonal. Recalling that $\widehat{\bm \Sigma} = N^{-1} \bm
X_c^{T} \bm X_c$, we have the eigendecomposition $\widehat{\bm \Sigma} = \bm U
\bm D \bm U^{T}$, where $\bm U$ is the matrix of eigenvectors of $\widehat{\bm
  \Sigma}$ and $\bm D = N^{-1} \bm \Delta$ is the diagonal matrix of eigenvalues
of $\widehat{\bm \Sigma}$. Now, we can obtain $\bm M$ and $\bm D$ efficiently
from the eigenvalue decomposition of $\bm X_c \bm X_c^{T} = \bm M \bm D \bm
M^{T}$. Hence, we eigendecomposition of $\bm X_c \bm X_c^{T} \in \mathbb{R}_{N
  \times N}$. Next, we compute $\bm U = \bm X_c^{T} \bm M \bm D^{+/2}$, where
$\bm D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$. We then determine $q$, the
number of numerically nonzero eigenvalues present in $\bm D$, by calculating the
number of eigenvalues that exceeds some tolerance value, say, $\epsilon = 1
\times 10^{-6}$. After calculating $q$, we extract $\bm U_1$ by retaining only
the first $q$ columns of $\bm U$.

>>>>

\citeauthor{Friedman:1989tm} proposed the popular regularized discriminant
analysis (RDA) classifier that utilizes a biased covariance-matrix estimator to
partially pool the sample covariance matrices from linear and quadratic
disriminant analysis before shrinking the resulting estimator towards a scaled
identity matrix. The RDA classifier's two tuning parameters are typically
estimated via a computationally burdensome cross-validation procedure from a
grid of candidate values. We reformulate the RDA classifier in the small-sample,
high-dimensional setting and then show that the classification decision rule is
equivalent to one in a subspace having a much lower dimension. As a result, the
dimension reduction involved yields a substantial reduction in computation
during model selection. Our parameterization also offers interpretability
previously lacking with the RDA classifier. We demonstrate with three artificial
and four real high-dimensional data sets that our proposed classifier is often
superior to several recently proposed sparse and regularized classifiers in
terms of classification accuracy. Finally, we provide an implementation of our
proposed classifier in the {\tt sparsediscrim} R package, available on CRAN.

@@

We reparameterize the \emph{RDA} classifier similar to \cite{Hastie:2008dt} and
\cite{Halbe:2007ea} and refer to our parameterization as the high-dimensional
\emph{RDA} (\emph{HDRDA}) classifier. First, we show the pooling parameter in
the \emph{HDRDA} classifier determines the contribution of each training
observation to the estimation of each class covariance matrix, enabling
interpretability that has been previously lacking with the \emph{RDA} classifier
\citep{Bensmail:1996fw}. Next, applying a reasoning similar to \cite{Ye:2006tq},
we establish that the matrix operations corresponding to the null space of the
pooled sample covariance matrix are redundant and can be discarded from the
\emph{HDRDA} decision rule without loss of classificatory information. Moreover,
because the \emph{HDRDA} decision rule is invariant to adjustments to the zero
eigenvalues, the decision rule in the original feature space is equivalent to a
decision rule in a much lower dimension (i.e., the intrinsic dimension), such
that the matrix inverses and determinants of relatively small matrices can be
rapidly computed. Thus, the dimension reduction inherent in the \emph{HDRDA}
classifier yields a substantial reduction in computation during model
selection. Also, as a result, we show that a large number of shrinkage methods
that are special cases of the \emph{HDRDA} classifier have no effect on the zero
eigenvalues of the covariance-matrix estimators. Such methods include work from
\cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and other methods studied by
\cite{Ramey:2013ji} and \cite{Xu:2009fl}.

@@

In discriminant analysis we wish to assign an unlabeled vector $\bm x \in
\mathbb{R}_{p \times 1}$ to one of $K$ unique, known classes by constructing a
classifier from $N$ training observations. Let $\bm x_i = (x_{i1}, \ldots,
x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots,
N)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$. We
assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution $p(\bm
x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$
is the probability density function of the $k$th class and $p(\omega_k)$ is the
prior probability of class membership of the $k$th class. We further assume that
each of $p(\omega_k)$ is equal, $k = 1, \ldots, K$.

@@

The smallest eigenvalues of $\widehat{\bm \Sigma}_k$ and the directions
associated with their eigenvectors can highly influence \eqref{eq:qda}. In fact,
the eigenvalues of $\widehat{ \bm \Sigma}_k$ are well-known to be biased if $p
\ge n_k$ such that the smallest eigenvalues are underestimated
\citep{Seber:2004uh}. Moreover, if $p > n_k$, then rank$(\widehat{\bm \Sigma}_k)
\le n_k$, which implies that at least $p - n_k$ eigenvalues of $\widehat{\bm
  \Sigma}_k$ are zero. Furthermore, although more feature information is
available to discriminate among the $K$ classes, if $p > n_k$, \eqref{eq:qda} is
incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does not exist.

@@

Here, we define the \emph{HDRDA} classifier by first formulating the
covariance-matrix estimator $\widehat{\bm \Sigma}_k(\lambda)$ and demonstrating
its clear interpretation as a linear combination of the crossproducts of the
training observations centered by their respective class sample means. We define
the convex combination
\begin{align}
  \widehat{\bm \Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm \Sigma}_k + \lambda
  \widehat{\bm \Sigma}, \quad k = 1, \ldots, K,\label{eq:sig-lambda-alternative}
\end{align}
where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. By rewriting
\eqref{eq:sig-lambda-alternative} in terms of the observations $ \bm x_i$ $(i =
1, \ldots, N)$, each centered by its class sample mean, we attain a clear
interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is,
\begin{align}
	\widehat{\bm \Sigma}_k(\lambda)
  &= \left( 1 - \lambda + \frac{\lambda n_k}{N} \right) \widehat{\bm \Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}}^K n_{k'} \widehat{\bm \Sigma}_{k'} \nonumber \\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N
  I(y_i = k) \bm x_i \bm x_i^{T} +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k)
  \bm x_i \bm x_i^{T} \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i^{T},\label{eq:sig-lambda-alternative2}
\end{align}
where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$. We
can interpret \eqref{eq:sig-lambda-alternative2} as a covariance-matrix
estimator that borrows from $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov} to
better estimate $ \bm \Sigma_k$. From \eqref{eq:sig-lambda-alternative2},
$\lambda$ weights the contribution of each of the $N$ observations in estimating
$\bm \Sigma_k$. Hence, for $0 < \lambda \le 1$, we estimate $\bm \Sigma_k$ with
$N$ rather than $n_k$ observations.\footnote{TODO: Add more to the
  interpretation here.}

Below, we will show that the pooling operation is advantageous in increasing the
rank of each $\widehat{\bm \Sigma}_k(\lambda)$ from rank$(\widehat{\bm
  \Sigma}_k)$ to rank$(\widehat{\bm \Sigma})$ for $0 < \lambda \le 1$. Notice
that if $\lambda = 0$, then the observations from the remaining $K - 1$ classes
do not contribute to the estimation of $\bm \Sigma_k$, corresponding to
$\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the weights
$c_{ik}(\lambda)$ in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$,
corresponding to $\widehat{\bm \Sigma}$. For brevity, when $\lambda = 1$, we
define $\bm X = [\sqrt{c_{1k}(1)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} \bm
  x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma} = \bm X^{T} \bm X$. Similarly,
for $\lambda = 0$, we define $\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1^{T}, \ldots,
  \sqrt{c_{Nk}(0)} \bm x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma}_k = \bm
X_k^{T} \bm X_k$. In Figure \ref{fig:hdrda-contours} we plot the contours of
five multivariate normal populations for $\lambda = 0$ with unequal covariance
matrices. As $\lambda$ approaches 1, the contours become more similar, resulting
in identical contours for $\lambda = 1$.

\begin{figure}
<<hdrda_contours, fig.width=8, fig.height=8, out.width='.33\\linewidth', fig.show='hold'>>=
rda_contours(lambda = 0)
rda_contours(lambda = 1/4)
rda_contours(lambda = 3/4)
rda_contours(lambda = 1)
@ 
\caption{Contours of five multivariate normal populations as a function of the
  pooling parameter $\lambda$.}
\label{fig:hdrda-contours}
\end{figure}

As we have discussed above, several eigenvalue adjustment methods have been
proposed to increase eigenvalues (approximately) equal to 0. To further improve
the estimation of $\bm \Sigma_k$ and to stabilize the estimator's inverse, we
define the eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
\begin{align}
	\tilde{\bm \Sigma}_k = \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:hdrda-cov}
\end{align}
where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
constant. Thus, the \emph{pooling} parameter $\lambda$ controls the amount
borrowed from $\widehat{\bm \Sigma}$ to estimate $\bm \Sigma_k$, and the
\emph{shrinkage} parameter $\gamma$ determines the degree of eigenvalue
shrinkage. The choice of $\alpha_k$ allows for a flexible formulation of
covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1, \ldots,
K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}. Similarly,
if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form similar to the
\emph{RDA} classifier from \cite{Friedman:1989tm}. Now, substituting
\eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the \emph{HDRDA} classifier
\begin{align}
	D_{HDRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)^{T}\tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k|. \label{eq:hdrda}
\end{align}
For $\gamma > 0$, $\tilde{\bm \Sigma}_k$ is nonsingular such that $\tilde{\bm
  \Sigma}_k^{-1}$ can be substituted for $\tilde{\bm \Sigma}_k^{+}$ in
\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\bm \Sigma}_k|$
equal to the product of the positive eigenvalues of $\tilde{\bm
  \Sigma}_k$. Following \cite{Friedman:1989tm}, we select $\lambda$ and $\gamma$
from a grid of candidate models via cross-validation \citep{Hastie:2008dt}. We
provide an implementation of \eqref{eq:hdrda} in the {\tt hdrda} function in the
{\tt sparsediscrim} R package, available on CRAN.

@@

To further improve the estimation of $\bm \Sigma_k$ and to stabilize the inverse
of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased
covariance-matrix estimator
\begin{align}
          \widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm
            \Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm
              \Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
          \end{align}
where $\gamma \in [0, 1]$ is a regularization parameter that controls the
shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
of the eigenvalues of \eqref{eq:sig-lambda}. Finally, the RDA classifier is
established by substituting \eqref{eq:sig-rda} into
\eqref{eq:qda}.\footnote{TODO: Mention the values of $\alpha_k$ and $\gamma$
  needed to resemble HDRDA.}

\subsubsection{Ridge-Like Estimators}

TODO: List a few estimators here along with $\alpha_k = 1$ such that HDRDA
includes them all. For example, set $\gamma = \text{trace}(\widehat{\bm
  \Sigma}_k) / \min(n_k, p)$ for Srivastava and Kubokawa.

\section{Properties of the HDRDA Classifier}
\label{sec:hdrda-properties}

@@

Thus, by incorporating each $\bm x_i$ into the estimation of $\bm \Sigma_k$, we
increase the rank of $\widehat{\bm \Sigma}_k(\lambda)$ to $q \approx N$ if
$\lambda \ne 0$. Next, we provide an essential result that enables us to prove
that \eqref{eq:hdrda} is invariant to adjustments to the eigenvalues of $\tilde{
  \bm \Sigma}_k$ corresponding to the null space of $\widehat{\bm \Sigma}$.

@@

\begin{proposition}\label{proposition:hdrda-W_k}
Let $\bm W_k$ be defined as above. Then, $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$
and
	\begin{align}
		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - \alpha_k(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T} \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},\label{eq:W_k_inv}
    \intertext{where}
    \bm Q_k &= \bm I_{n_k} + \alpha_k(1 - \lambda) \bm X_k \bm U_1 \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T}\label{eq:Q_k}
    \intertext{and}
    \bm \Gamma_k &= \alpha_k \lambda \bm D_q + \gamma \bm I_q\label{eq:Gamma_k}.
	\end{align}
\end{proposition}
\begin{proof}
To calculate $|\bm W_k|$, we apply Theorem 18.1.1 from \cite{Harville:2008wja},
which states that $|\bm A + \bm B \bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} +
\bm C \bm A^{-1} \bm B|$. Thus, setting $\bm A = \bm \Gamma_k$, $\bm B =
\alpha_k (1 - \lambda) \bm U_1^{T} \bm X_k^{T}$, $\bm T = \bm I_{n_k}$, and $\bm
C = \bm X_k \bm U_1$, we have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$. Also,
\eqref{eq:W_k_inv} follows from the well-known Sherman-Woodbury formula
\citep[Theorem 18.2.8]{Harville:2008wja}.
\end{proof}

@@

        Compute $(\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x -
        \xbar_k) + \log | \bm W_k |$ for each row $\bm x$ of $\bm X_{test}$
      }
      Classify test observations $\bm X_{test}$ using \eqref{eq:hdrda-decomposed}

@@

A primary contributing factor to the efficiency of Algorithm
\ref{alg:hdrda_model_selection} is our usage of the compact singular value
decomposition (SVD) (sometimes called the reduced or fast SVD). Rather than
computing the eigenvalue decomposition of $\widehat{\bm \Sigma}$ to obtain $\bm
U_1$, we instead obtain $\bm U_1$ by computing the eigendecomposition of a much
smaller $N \times N$ matrix when $p \gg N$
\cite[Chapter~18.3.5]{Hastie:2008dt}. Applying the SVD, we decompose $\bm X_c =
\bm M \bm \Delta \bm U^{T}$, where $\bm M \in \mathbb{R}_{N \times p}$ is
orthogonal, $\bm \Delta \in \mathbb{R}_{p \times p}^{\ge}$ is a diagonal matrix
consisting of the singular values of $\bm X_c$, and $\bm U \in \mathbb{R}_{p
  \times p}$ is orthogonal. Recalling that $\widehat{\bm \Sigma} = N^{-1} \bm
X_c^{T} \bm X_c$, we have the eigendecomposition $\widehat{\bm \Sigma} = \bm U
\bm D \bm U^{T}$, where $\bm U$ is the matrix of eigenvectors of $\widehat{\bm
  \Sigma}$ and $\bm D = N^{-1} \bm \Delta$ is the diagonal matrix of eigenvalues
of $\widehat{\bm \Sigma}$. Now, we can obtain $\bm M$ and $\bm D$ efficiently
from the eigenvalue decomposition of the $N \times N$ matrix $\bm X_c \bm
X_c^{T} = \bm M \bm D \bm M^{T}$. Next, we compute $\bm U = \bm X_c^{T} \bm M
\bm D^{+/2}$, where $\bm D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$. We then
determine $q$, the number of numerically nonzero eigenvalues present in $\bm D$,
by calculating the number of eigenvalues that exceeds some tolerance value, say,
$\epsilon = 1 \times 10^{-6}$. After calculating $q$, we extract $\bm U_1$ by
retaining only the first $q$ columns of $\bm U$.
